---
title: "Kansas Judicial Retention Election"
author: "Earl F Glynn<br><small>watchdoglab.substack.com</small>"    
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

Create this notebook from 2010/2016 scripts.

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r Constants}
FILE_PREFIX   <- "2022-11-General"    # no spaces
```

```{r}
EXCEL_LIMIT <- 2^20
```

```{r}
county.abbr <- c(
  "AL","AN","AT","BA","BT","BB","BR","BU","CS","CQ","CK","CN","CA","CY",
  "CD","CF","CM","CL","CR","DC","DK","DP","DG","ED","EK","EL","EW","FI",
  "FO","FR","GE","GO","GH","GT","GY","GL","GW","HM","HP","HV","HS","HG",
  "JA","JF","JW","JO","KE","KM","KW","LB","LE","LV","LC","LN","LG","LY",
  "MP","MN","MS","ME","MI","MC","MG","MR","MT","NM","NO","NS","NT","OS",
  "OB","OT","PN","PL","PT","PR","RA","RN","RP","RC","RL","RO","RH","RS",
  "SA","SC","SG","SW","SN","SD","SH","SM","SF","ST","SV","SU","TH","TR",
  "WB","WA","WS","WH","WL","WO","WY")          
```

## Packages

```{r Packages}
library(tidyverse)
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

 Graphics

```{r Graphics}
library(maps)
```

## Helper functions

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       |>
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  |>
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       |>
  scroll_box(height = height, width = width) #  |>
# as.character()                             |>      # Jupyter Only
# display_html()                                      # Jupyter Only
}
```

# Read election results

Source is [2022 Kansas General Election Unoffical Results page](https://ent.sos.ks.gov/kssos_ent.html), Last and FINAL Update: 11/14/2022 6:18 PM

County results screen scraped from SOS site.   

```{r}
results <-
  read_xlsx("../County-Results-ent.sos.ks.gov-2022-11-14-1818-FINAL.xlsx",
            guess_max = EXCEL_LIMIT)                    |>    #####
  filter(str_detect(Contest, "Supreme Court Justice"))  |>
  separate(Candidate, c("Name", "Retain"), sep = "-")   |>
  select(-Contest,
         -starts_with("State"), 
         -CountyReport, 
         -`County %`, 
         -Precincts,
         Votes = `County Votes`)                   |>
  mutate(Retain = str_replace_all(Retain, '"', "") |>
                  str_trim(),
         Name   = str_trim(Name))                  |>
  group_by(County, Name, Retain)                   |>
  pivot_wider(names_from = "Retain",
              values_from = Votes)                 |>
  ungroup()                                        |>
  mutate(Retain = YES > NO)

dim(results)
```          

```{r}
results |> Show(height = "400px")
```     


```{r}
write_xlsx(results, 
          paste0(FILE_PREFIX, "-County-Results-Judical-Retention.xlsx"))
```

```{r}
countiesVotingNotToRetain <- 
  results |> 
    filter(!Retain)  

dim(countiesVotingNotToRetain)
```

```{r}
countiesVotingNotToRetain |>  Show(height = "400px")
```


```{r}
write_xlsx(countiesVotingNotToRetain, 
          paste0(FILE_PREFIX, "-Counties-Voting-Against-Judical-Retention.xlsx"))
```

```{r}
justices <- countiesVotingNotToRetain$Name |> unique() |> sort()
justices
```

```{r}
countiesVotingNotToRetain                   |>
  summarize(n         = n(),
            nCounty   = n_distinct(County),
            nJudtices = n_distinct(Name))    |>
  Show()
```

# Example of defining map labels and colors

Someday rework this old code originally used in 2010 and 2016

```{r}
justiceNoCounties <- 
  countiesVotingNotToRetain  |>
  filter(Name == justices[1])

justiceNoCounties |> Show()
```

## Get list of county names and abbreviations 

in FIPS order by county name

```{r}
m <- map('county', 'kansas', plot=FALSE)
countyNames <- str_to_title( str_replace(m$names, "kansas,", "") )
countyNames[57] <- "McPherson"   # Force this fix  for other matches 
```

```{r}
countyNames
```
```{r}
county.abbr
```

## Set default `Color` and `Labels` for each county

```{r}
Color  <- rep("white", 105)
Labels <- rep("", 105)
```

```{r}
notRetained <- countyNames %in% justiceNoCounties$County
table(notRetained)
```

```{r}
which(notRetained) 
```

```{r}
Labels[notRetained] <- county.abbr[notRetained]
Labels
```

```{r}
Color[notRetained] <- rep("blue", sum(notRetained))
Color
```


```{r example}
m <- map('county', 'kansas', fill = TRUE, col = Color)
  map.text("county","kansas",
    labels = Labels, cex=1.15, add=TRUE, col="white")
```

# Map helper function

```{r}
plotNoCounties <- function (justiceNoCounties,
                            MapTitle="Counties Voted NOT to Retain")
{
  m <- map('county', 'kansas', plot=FALSE)
  countyNames <- str_to_title( str_replace(m$names, "kansas,", "") )
  countyNames[57] <- "McPherson"   # Force this fix  for other matches
  
  Color  <- rep("white", 105)
  Labels <- rep("", 105)
  
  notRetained <- countyNames %in% justiceNoCounties$County
  
  Color[notRetained]  <- rep("blue", sum(notRetained))
  Labels[notRetained] <- county.abbr[notRetained]
  
  m <- map('county', 'kansas', fill = TRUE, col = Color)
  
  map.text("county","kansas",
           labels = Labels, cex=1.15, 
           add=TRUE, col = "white")
  mtext(paste(sum(notRetained), MapTitle, justiceNoCounties$Name[1]), cex=1.5)

  mtext(paste0("Source:  Kansas Secretary of State, Nov 8, 2022 General Election\n",  #####
               "Final Unofficial Online Election Results, Nov 14, 2022"),             #####
    BOTTOM<-1, adj=0, line=0.5, cex=0.75)
  
  mtext("watchdogLab.substack.com", 1, adj=1, line=0.5, col="blue")                            
}                        
```


# Map for each justice losing one or more counties

```{r justice}
for (iJustice in 1:length(justices))
{
  plotNoCounties(countiesVotingNotToRetain  |>
               filter(Name == justices[iJustice]))     
}
```


# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

