---
title: "Dennis Pyle - 2022 Gubernatorial Spoiler Map"
author: "Earl F Glynn<br>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
FILE_PREFIX <- "2022-11-General"                                           #####
```

```{r Constants}
SOURCE <- "Kansas Secretary of State's Unofficial Election Results, Nov. 14, 2022 Final"   #####
stamp  <- paste("watchdoglab.substack.com", format(Sys.time(), "%Y-%m-%d"))
```

```{r}
EXCEL_LIMIT <- 2^20
```

## Packages

```{r Packages}
library(tidyverse)
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

 Graphics

```{r Graphics}
library(scales)      # comma, comma_format
```

## ggplot theme

[ggplot Themes Gallery](https://www.datanovia.com/en/blog/ggplot-themes-gallery/)

```{r ggplotTheme}
theme_set(theme_minimal() +

          theme(axis.text             = element_text(size = 10),
                axis.title            = element_text(size = 14),

                plot.caption          = element_text(size = 10,
                                                     hjust = c(0.0,1.0)),
                plot.caption.position = "plot",   # R 4.0.0

                plot.title.position   = "plot",
                title                 = element_text(size = 14))
           )

COLOR_BAR     <- "skyblue"
COLOR_OUTLINE <- "grey80"
```

## Helper functions

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       %>%
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  %>%
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       %>%
  scroll_box(height = height, width = width) #  %>%
# as.character()                             %>%      # Jupyter Only
# display_html()                                      # Jupyter Only
}
```

# Read 2022-11 Results

Source is [2022 Kansas General Election Unoffical Results page](https://ent.sos.ks.gov/kssos_ent.html), Last and FINAL Update: 11/14/2022 6:18 PM

Statewide results

```{r}
#                   Kelly   Schmidt  Cordell  Pyle
statewideVotes <- c(492209, 471323,  10888,   20057)                               #####
```

```{r}
PyleStatePercent <- round(100 * statewideVotes[4] / sum(statewideVotes), 2)
PyleStatePercent
```

County results screen scraped from SOS site.

```{r}
results <- 
  read_xlsx("../County-Results-ent.sos.ks.gov-2022-11-14-1818-FINAL.xlsx")  %>%    #####
  filter(Contest == "Governor / Lt. Governor")  %>%
  select(-Contest, -starts_with("State"), -CountyReport) %>%
  separate(Candidate, c("Party", "CandidateName"), sep = "-")

dim(results)
```
```{r}
results %>% Show(height = "400px")
```
Let's compute total votes by county

```{r}
countyTotals <- 
  results           %>%
  group_by(County)  %>%
  summarize(TotalVotes = sum(`County Votes`))  %>%
  ungroup()

dim(countyTotals)
```

Let's recompute candidate percentages for more precision

```{r}
results <- 
  results                                  %>%
  inner_join(countyTotals, by = "County")  %>%
  mutate(Percent = round(100 * `County Votes` / TotalVotes, 2))


results %>%  Show(height = "400px")
```

```{r}
write_xlsx(results, paste0(FILE_PREFIX, "-Results-for-Gubernatorial-Candidates.xlsx"))
```

```{r}
Pyle <- 
  results                                %>% 
  filter(CandidateName == "Dennis Pyle") %>%
  select(County, `County Votes`, ColorIndex = `County %`,  Percent)

dim(Pyle)
```

```{r}
Pyle %>% Show(height = "400px")
```

Make sure SOS rounding (now ColorIndex) consistent with computed percentages

```{r}
all(round(Pyle$Percent) == Pyle$ColorIndex)
```
```{r}
countyCounts <- 
  as.data.frame( table(Pyle$ColorIndex) )  %>%
  rename(PercentInterval = Var1)

countyCounts %>%  Show()
```

```{r}
sum(countyCounts$Freq)
```


```{r}
write_xlsx(Pyle, paste0(FILE_PREFIX, "-County-Results-for-Pyle.xlsx"))
```

```{r}
voteQuantiles <- quantile(Pyle$Percent,
                          c(0.0, 0.05, 0.10, 0.25, 0.50,
                             0.75, 0.90, 0.95, 1.0))
voteQuantiles
```

# Density plot of Pyle's County Percentage

```{r DensityPlot, fig.width = 8, fig.height = 6}
ggplot(Pyle, aes(x = Percent, y=..density..))                       +
geom_histogram(fill = "skyblue", color = "grey80")                  +
geom_density()                                                      +
scale_x_continuous(breaks = 0:10)                                   +
labs(title =  "Density Plot of Dennis Pyle's County Percentages in Governor's Race",
     subtitle = "Variation across the 105 Kansas Counties in Nov. 2022 General Election",
     x = "Dennis Pyle's County Vote Percentage",
     y = "Density",
     caption = c(SOURCE, stamp))                                    +
theme(axis.ticks.x       = element_line(),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank())
```

# Map Pyle Results

```{r}
library(maps)
library(RColorBrewer) # brewer.pal
```

## Palette setup

```{r}
 brewer.pal.info["Greys",]  %>%  Show()
```

```{r, fig.height = 1}
par(mar=c(1,1,1,1))
display.brewer.pal(9,"Greys")
```
Set complementary background color for text

```{r}
color.map  <- brewer.pal(9, "Greys")
color.text <- c(rep("black", 4), rep("white", 4), "yellow")
```

## Put data in FIPS order

McPherson is first under M's in FIPS order

```{r}
CapWords <- function(s, strict = FALSE)
{
    cap <- function(s) paste(toupper(substring(s,1,1)),
                  {s <- substring(s,2); if(strict) tolower(s) else s},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}
```

```{r}
county.abbr <- c(
  "AL","AN","AT","BA","BT","BB","BR","BU","CS","CQ","CK","CN","CA","CY",
  "CD","CF","CM","CL","CR","DC","DK","DP","DG","ED","EK","EL","EW","FI",
  "FO","FR","GE","GO","GH","GT","GY","GL","GW","HM","HP","HV","HS","HG",
  "JA","JF","JW","JO","KE","KM","KW","LB","LE","LV","LC","LN","LG","LY",
  "MP","MN","MS","ME","MI","MC","MG","MR","MT","NM","NO","NS","NT","OS",
  "OB","OT","PN","PL","PT","PR","RA","RN","RP","RC","RL","RO","RH","RS",
  "SA","SC","SG","SW","SN","SD","SH","SM","SF","ST","SV","SU","TH","TR",
  "WB","WA","WS","WH","WL","WO","WY")
```

```{r}
m <- map('county', 'kansas', plot=FALSE)
names(m)
```

```{r}
countyNames <- tibble(County = CapWords(sub("kansas,","", m$names)))
countyNames$County[57] <- "McPherson"   # Force this fix
countyNames$County
```

```{r}
Pyle <- 
  countyNames  %>%
  inner_join(Pyle, by = "County")

Pyle$County
```

## Plot map

```{r, fig.width = 10, fig.height = 7.5}
layout(rbind(1,2), heights=c(9,1))

par(mar=c(0,1,1,1))

county.label <- paste(county.abbr, "\n", 
                      sprintf("%.2f", Pyle$Percent), "%", sep="")

m <- map('county', 'kansas', fill = TRUE, col = color.map[Pyle$ColorIndex])

map.text("county", "kansas",
  col = color.text[Pyle$ColorIndex],
  labels = county.label, cex = 0.8, add = TRUE)

mtext(expression(bold("Dennis Pyle Gubernatorial Voters [%]")), cex = 2)

par(mar=c(2, 6, 0, 6))
barplot(rep(1,length(color.map)), col=color.map, space=0.0, border="black", axes=FALSE)
text(1:length(color.map) - 0.5, rep(0.5,length(color.map)),
     paste0(1:9, "%"),
     cex=1,
     col=color.text)
mtext(SOURCE, BOTTOM <- 1, adj = 0, line = 0.5, cex = 0.8, col = "blue")
mtext(stamp,  BOTTOM,      adj=1,   line=0.5,   cex = 0.8, col = "blue")

mtext(paste0("Statewide Dennis Pyle Voters: ",
             sprintf("%.2f", PyleStatePercent), "%"),
      TOP<-3, line = 1, cex = 2)
```



# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

