---
title: "Screen Scrape Kansas Unofficial Election Results -- County Pages -- Exploration"
author: "Earl F Glynn<br>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

Starting point: https://ent.sos.ks.gov/kssos_ent.html

Example county link (JoCo):  https://ent.sos.ks.gov/johnson.html

Adapted from R scripts from 2010 and 2016 that scraped the 105 county pages and extracted election results.

Now use local download to store all results locally:  `Fetch-SOS-Unofficial-Results.bat`

`wget -e robots=off -N -t 50 -o wget.log -l 8 -r "https://ent.sos.ks.gov/kssos_ent.html"`   

Results stored in directory:

    ent.sos.ks.gov
    ent.sos.ks.gov/images
    
Files:

* `constitutional.html`

* `daysAfterElection.html`

* `favorite-races.html`

* `judicial.html`

* `ks-boe.html`

* `ks-senate.html`

* `kssos_ent.html`

* `Uncontested.html`

* For each county:  county.html 


```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
KansasCounties <- c(      # Names of 105 Kansas counties
  "Allen",      "Anderson",     "Atchison",    "Barber",    "Barton",
  "Bourbon",    "Brown",        "Butler",      "Chase",     "Chautauqua",
  "Cherokee",   "Cheyenne",     "Clark",       "Clay",      "Cloud",
  "Coffey",     "Comanche",     "Cowley",      "Crawford",  "Decatur",
  "Dickinson",  "Doniphan",     "Douglas",     "Edwards",   "Elk",
  "Ellis",      "Ellsworth",    "Finney",      "Ford",      "Franklin",
  "Geary",      "Gove",         "Graham",      "Grant",     "Gray",
  "Greeley",    "Greenwood",    "Hamilton",    "Harper",    "Harvey",
  "Haskell",    "Hodgeman",     "Jackson",     "Jefferson", "Jewell",
  "Johnson",    "Kearny",       "Kingman",     "Kiowa",     "Labette",
  "Lane",       "Leavenworth",  "Lincoln",     "Linn",      "Logan",
  "Lyon",       "Marion",       "Marshall",    "McPherson", "Meade",
  "Miami",      "Mitchell",     "Montgomery",  "Morris",    "Morton",
  "Nemaha",     "Neosho",       "Ness",        "Norton",    "Osage",
  "Osborne",    "Ottawa",       "Pawnee",      "Phillips",  "Pottawatomie",
  "Pratt",      "Rawlins",      "Reno",        "Republic",  "Rice",
  "Riley",      "Rooks",        "Rush",        "Russell",   "Saline",
  "Scott",      "Sedgwick",     "Seward",      "Shawnee",   "Sheridan",
  "Sherman",    "Smith",        "Stafford",    "Stanton",   "Stevens",
  "Sumner",     "Thomas",       "Trego",       "Wabaunsee", "Wallace",
  "Washington", "Wichita",      "Wilson",      "Woodson",   "Wyandotte")
```
                                                                                   

## Packages

```{r Packages}
library(tidyverse)
library(lubridate)
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

Scraping

```{r Scraping}
library(rvest)       # scraping
```


## Helper functions

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       %>%
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  %>%
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       %>%
  scroll_box(height = height, width = width) #  %>%
# as.character()                             %>%      # Jupyter Only
# display_html()                                      # Jupyter Only
}
```

# Exploration

```{r}
BASE <- "ent.sos.ks.gov-2022-11-14-1818-FINAL"                             #####
```

## Explore County

```{r}
  i <- 1
  county <- KansasCounties[i]
  filename <- paste0(BASE, "/", county, ".html")
  filename
```

```{r}
webPage <- read_html(filename)
```

```{r}
countyName <- 
  webPage                     %>%
  html_nodes("div.col-md-8")  %>%
  html_text()                 %>%
  str_trim()

countyName
```
```{r}
countyName <- (str_split(countyName, " ") %>% unlist())[3]
countyName
```

## Last Update

```{r}
updateTimes <-
  webPage                     %>%
  html_nodes("div.col-md-4")  %>%
  html_nodes("p")             %>%
  html_text()                 %>%
  str_trim()

updateTimes
```

```{r}
updateTimes <- str_sub(updateTimes, 2+str_locate(updateTimes, ":")[,1], str_length(updateTimes))
updateTimes
```
Use ISO 8601 format

```{r}
updateTimes <- parse_date_time(updateTimes, "%m/%d%y %H:%M Op") %>% str_sub(1, -4)
updateTimes
```

## Races

```{r}
races <- 
  webPage   %>%
  #html_nodes(xpath = "//div[@class = 'race']")   # XPath
  html_nodes("div.race")                          # CSS selector

length(races)
```

```{r}
races %>%
  html_nodes("h3")  %>%
  html_text()  
```

```{r}
index <- 3
```


```{r}
contest <- 
  races [[index]]  %>% 
    html_nodes("div.race-heading")  %>%
    html_nodes("h3")  %>%
    html_text()

contest
```

```{r}
reporting <- 
  races [[index]]  %>% 
    html_nodes("div.race-subheading")  %>%
    html_nodes("div.col-4")            %>%
    html_nodes("p")  %>%
    html_text()

reporting 
```

```{r}
precincts <- (str_split(reporting, ":") %>% unlist()  %>% str_trim)[2]
precincts
```

```{r}
races [[index]]  %>% 
  html_nodes("div.race-subheading")  %>%
  html_nodes("div.col-2")            %>%
  html_nodes("p")  %>%
  html_text
```

```{r}
tableHeader <-
  races [[index]]  %>% 
    html_nodes("div.race-subheading")  %>%
    html_nodes("p")  %>%
    html_text()

tableHeader
```

```{r}
candidates <- 
  races [[index]]                       %>% 
  html_nodes("div.race-candidate-list") %>%
  html_nodes("div.row")                

candidates
```

```{r}
candidateNames <- 
  candidates                 %>%
    html_nodes("div.col-4")  %>%
    html_text()              %>%
    str_trim()

candidateNames
```

```{r}
results <- 
  candidates                   %>%
    html_nodes("div.col-2")    %>%
    html_text()                %>%
    str_trim()                 %>%
    str_replace_all(",|%", "") %>%  # numbers not strings
    as.numeric()               %>%
    matrix(nrow = length(candidateNames), 
           byrow = TRUE,
           dimnames = list(1:length(candidateNames),
                           tableHeader[-1]))

results %>% Show()
```

```{r}
results <- bind_cols(County = countyName,
                      Contest   = contest,
                      Candidate = candidateNames,
                      Precincts = precincts,
                      results,
                      CountyReport = updateTimes[2],
                      StateReport  = updateTimes[1])

results %>% Show()
```

```{r}
# Scraping Functions

ExtractContest <- function(countyName, updateTimes, races, index)
{
    contest <-
    races [[index]]  %>%
      html_nodes("div.race-heading")  %>%
      html_nodes("h3")  %>%
      html_text()

  reporting <-
    races [[index]]  %>%
      html_nodes("div.race-subheading")  %>%
      html_nodes("div.col-4")            %>%
      html_nodes("p")  %>%
      html_text()

  precincts <- (str_split(reporting, ":") %>% unlist()  %>% str_trim)[2]

  tableHeader <-
    races [[index]]  %>%
      html_nodes("div.race-subheading")  %>%
      html_nodes("p")  %>%
      html_text()

  candidates <-
    races [[index]]                       %>%
    html_nodes("div.race-candidate-list") %>%
    html_nodes("div.row")

  candidateNames <-
    candidates                 %>%
      html_nodes("div.col-4")  %>%
      html_text()              %>%
      str_trim()

  results <-
    candidates                   %>%
      html_nodes("div.col-2")    %>%
      html_text()                %>%
      str_trim()                 %>%
      str_replace_all(",|%", "") %>%  # numbers not strings
      as.numeric()               %>%
      matrix(nrow = length(candidateNames),
             byrow = TRUE,
             dimnames = list(1:length(candidateNames),
                             tableHeader[-1]))

  contestResults <- bind_cols(County = countyName,
                              Contest   = contest,
                              Candidate = candidateNames,
                              Precincts = precincts,
                              results,
                              CountyReport = updateTimes[2],
                              StateReport  = updateTimes[1])

  return(contestResults)
}


ExtractCounty <- function(BASE, county)
{
  # County filename

  filename <- paste0(BASE, "/", county, ".html")
  webPage <- read_html(filename)

  # Extracted county name
  countyName <-
    webPage                     %>%
    html_nodes("div.col-md-8")  %>%
    html_text()                 %>%
    str_trim()

  countyName <- (str_split(countyName, " ") %>% unlist())[3]

  # Update Times
  updateTimes <-
    webPage                     %>%
    html_nodes("div.col-md-4")  %>%
    html_nodes("p")             %>%
    html_text()                 %>%
    str_trim()

  updateTimes <- str_sub(updateTimes,
                         2 + str_locate(updateTimes, ":")[,1],
                         str_length(updateTimes))
  updateTimes <-
    parse_date_time(updateTimes,"%m/%d%y %H:%M Op") %>%  # ISO 8601
    str_sub(1, -4)

  ## Races
  races <-
    webPage   %>%
    #html_nodes(xpath = "//div[@class = 'race']")   # XPath
    html_nodes("div.race")                          # CSS selector

  races %>%
    html_nodes("h3")  %>%
    html_text()

  countySummary <- NULL
  for (index in 1:length(races))
  {
    contestResults <- ExtractContest(countyName, updateTimes, races, index)
    countySummary <- bind_rows(countySummary, contestResults)
  }

  return(countySummary)
}
```

```{r}
# Process Counties

stateSummary <- NULL

for (i in 1:length(KansasCounties))
{
  cat(i, KansasCounties[i], "\n")
  countySummary <- ExtractCounty(BASE, KansasCounties[i])
  
  stateSummary <- bind_rows(stateSummary,
                            countySummary)
}

nrow(stateSummary)
```

```{r}
write_xlsx(stateSummary, paste0("County-Results-", BASE, ".xlsx"))
```

# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

